Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-21 papers/FX/draft.tex
Finished the proof with the 1st power-utility function.

\input{../../physics_common}
\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

\subsection{power-law utility function}
In this section we follow Routledge and Zin
\cite{routledge2010generalized} and assume 
that the utility function of consumption $C$ is a power function%
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r$ be the return on the
riskless state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation}

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation*}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation*}
Assume $\alpha, \beta > 1$, $0 < p < 1$, $K > 0$. Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + (1 + b) \E [u(\delta v) - u(C); C < \delta v]
\]
\begin{theorem}
  Let
  \[
  G(\alpha, \beta, K, K') =
  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
  \]
  $G(\alpha, \beta, K, K')$ is monotonically increasing with
  $\alpha$.
\end{theorem}

@@

  \item When $C < \delta v$ implies $X < -q$ for some $q > 0$, we have

@@

      \pd{\mathcal U(F)}{\alpha}

@@

        (1 - \phi) e^r + \phi e^x

@@

      \underbrace{

@@

        (1 - \phi) e^r + \phi e^{-x}

@@

      }_{L_1}\\

@@

      }_{L_2} \\

@@

        \xi (1 - \phi)^\xi e^{r \xi}

@@

        \xi (\alpha - 1) (1 - \phi)^\xi e^{r \xi}

@@

        (1 - \phi) e^r + \phi e^{-x}
      \alpha K^{\alpha} (K + x)^{-\alpha - 1} dx \\
      &=& {\alpha L_2 \over c} \\
      L_1 &=& c \E[Y \ln(Y/K)]

@@

        x \ln\left({x \over K}\right)
      \right] = {1 \over x} > 0

@@

    That is, $x \ln(x/K)$ defined on $(0, \infty)$ is a convex

@@

      {L_1 \over c} = \E [Y \ln(Y/K)]

@@

      \E Y \times \ln(\E Y/K)

@@

      {\alpha L_2  \over c K}

@@

  \end{enumerate}

@@

@@

        (1 - \phi) e^r + \phi

@@

        p [(K + q)^{-\alpha} b + K^{-\alpha}]

@@

        \alpha \xi [(1 - \phi) e^r + \phi]^\xi

@@

\end{proof}

%% \[
%% F(x) = \left\{
%% \begin{array}{ll}
%%     p M^\alpha / |x|^\alpha & \text{ if } x \leq -M \\
%%     F_u(x) & \text{ if } -M < x < M\\
%%     1 - p M^\alpha x^\alpha & \text{ if } x \geq M
%% \end{array}
%% \right.
%% \]
%% where the distribution function $F_u(\cdot)$ satisfies
%% \begin{eqnarray*}
%%   F_u(-M) &=& p \\
%%   F_u(M) &=& 1 - p
%% \end{eqnarray*}
%% In this case we have
%% \begin{eqnarray*}
%%   && u(v) + b u(\delta v) \\
%%   &=&
%%   {
%%     (1 + b) p \alpha M^\alpha
%%     \over
%%     \kappa
%%   }
%%   \int_{-\infty}^{-M}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} |x|^{-\kappa - 1}
%%   dx + \\
%%   &&
%%   (1 + b) \int_{-M}^{\delta v}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} d F_u(x) + \\
%%   &&
%%   {
%%     (1 + b) p \alpha M^\alpha
%%     \over
%%     \kappa
%%   }
%%   \int_{M}^{\infty}
%%   ((1 - \phi) e^r + \phi e^x)^{-\kappa} x^{-\kappa - 1}
%%   dx \\
%%   &=&
%%   I_1 + I_2 + I_3
%% \end{eqnarray*}

@@

\bibliographystyle{unsrt}
\bibliography{../../thesis/econophysics}
\end{document}

>>>>

\usepackage{amssymb,amsmath,amsthm,amsfonts}
\usepackage{mathrsfs}
\usepackage{dsfont}
\usepackage{enumerate}

%% \newtheorem{mdef}{Definition}

\newcommand{\eqsplit}[2]{
  \begin{equation}\label{#2}
    \begin{split}
      #1
    \end{split}
  \end{equation}}
\newcommand{\eqnsplit}[1]{
  \begin{eqnarray*}
    #1
  \end{eqnarray*}}
\newcommand{\tran}[1]{
  \tilde{#1}
}
\newcommand{\td}[2]{
  \frac{d #1}{d #2}
}
\newcommand{\pd}[2]{
  \frac{\partial #1}{\partial #2}
}
\newcommand{\ppd}[2]{
  \frac{\partial^2 #1}{\partial #2^2}
}
\newcommand{\pdd}[3]{
  \frac{\partial^2 #1}{\partial #2 \partial #3}
}
\newcommand{\otd}[1]{
  \frac{d}{d #1}
}
\newcommand{\opd}[1]{
  \frac{\partial}{\partial #1}
}
\newcommand{\oppd}[1]{
  \frac{\partial^2}{\partial #1^2}
}
\newcommand{\opdd}[2]{
  \frac{\partial^2}{\partial #1 \partial #2}
}
\newcommand{\ket}[1]{
  |#1\rangle
}
\newcommand{\bra}[1]{
  \langle#1|
}
\newcommand{\inn}[1]{
  \langle#1\rangle
}
\newcommand{\mean}[1]{
  \langle#1\rangle
}
\newcommand{\tr}{
  \text{tr}\,
}
\newcommand{\re}{
  \text{Re}\,
}
\newcommand\im{
  \text{Im}\,
}
\newcommand{\arcsinh}{
  \sinh^{-1}
}
\newcommand{\arccosh}{
  \cosh^{-1}
}
\newcommand{\erfc}{
  \text{erfc}
}
\newcommand{\E}{
  \mathbb{E}
}
\renewcommand{\P}{
  \mathbb{P}
}
\newcommand{\I}{
  \mathbf{1}
}
\newcommand{\1}[1]{
  \mathbf{1}_{\{#1\}}
}
\newcommand{\diag}{
  \text{diag\,}
}
\newcommand{\pop}[1]{
\varepsilon_{#1}
}

\newcommand{\M}{
  {\text{max}}
}
\newcommand{\m}{
  {\text{min}}
}
% \newcommand{\ph}{
%   {\text{arg}\,}
% }
\newcommand\erf{
  \text{erf}
}
\renewcommand\vec[1]{
  \mathbf{#1}
}
\newcommand\cov{
  \text{cov}
}
\newcommand\var{
  \text{var}
}
\newcommand\cor{
  \text{cor}
}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\newcommand\mtx[1]{
  \mathbf{#1}
}
\renewcommand \L {
  \mathfrak{leb}
}
\newcommand{\hot}{
  \text{h.o.t.}
}

%% \newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}{Acknowledgement}
\newtheorem{algorithm}{Algorithm}
\newtheorem{axiom}{Axiom}
\newtheorem{case}{Case}
\newtheorem{claim}{Claim}
\newtheorem{conclusion}{Conclusion}
\newtheorem{condition}{Condition}
\newtheorem{conjecture}{Conjecture}
\newtheorem{corollary}{Corollary}
\newtheorem{criterion}{Criterion}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{exercise}{Exercise}
\newtheorem{lemma}{Lemma}
\newtheorem{notation}{Notation}
\newtheorem{problem}{Problem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{solution}{Solution}
\newtheorem{summary}{Summary}

\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

\subsection{power-law utility function}
In this section we follow Routledge and Zin
\cite{routledge2010generalized} and assume 
that the utility function of consumption $C$ is a power function%
\begin{equation}
  \label{eq:power_utility}
  u(C)=\frac{1}{1-\gamma}C^{1-\gamma}%  
\end{equation}
where $\gamma > 0$. For simplicity we assume the investor
initially has one unit of wealth. Let $r > 0$ be the return on the
risk free state bond and $X$ be the return on equity. Portfolio weights
are $\left(1-\phi\right)$ and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  %% u(C)=\frac{1}{1-\gamma}V^{1-\gamma} \max \left\{
  %% 0, 1 + (1-\phi) r  +\phi X
  %% \right\}^{1-\gamma} \label{3}%
  C = (1 - \phi) e^r + \phi e^X
\end{equation}

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation*}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation*}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and
\begin{equation}
  \label{eq:xxie1.1}
  0 < K <
  {\alpha \over \alpha + 1}
  (e^r - 1)^\xi
  \exp\left(
    - r \xi
    + {1 \over \alpha}
  \right)
\end{equation}
Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + b \E [u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v]
\]
\begin{theorem}
  Let
  \[
  G(\alpha, \beta, K, K') =
  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
  \]
  $G(\alpha, \beta, K, K')$ is monotonically increasing with
  $\alpha$.
\end{theorem}

@@

  \item When $C < \delta v$ implies $X < -q$ for some $q \geq 0$, we have

@@

      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\

@@

            (1 - \hat\phi) e^r + \hat\phi e^x

@@

      &=& \underbrace{

@@

          (1 - \hat\phi) e^r + \hat\phi e^{-x}

@@

      }_{L_1} \\

@@

      }_{L_2}

@@

        \xi (1 - \hat\phi)^\xi e^{r \xi}

@@

        \xi (\alpha - 1) (1 - \hat\phi)^\xi e^{r \xi}

@@

        (1 - \hat\phi) e^r + \hat\phi e^{-x}
      \right]^{-\xi} {
        \alpha K^{\alpha} x dx
      &=& {\alpha L_2 \over c} - K \\
      L_1 &=& c \E \left[
        (Y + K) \ln \left(1 + {Y \over K} \right)

@@

      (x + K) \ln\left(1 + {x \over K}\right)
    \right] = {1 \over x + K} > 0

@@

    That is, $(x + K) \ln(1 + x/K)$ defined on $(0, \infty)$ is a convex

@@

    {L_1 \over c} = \E \left[
      (Y + K) \ln \left(1 + {Y \over K} \right)

@@

    (\E Y + K) \times \ln(1 + \E Y/K)

@@

@@

@@

    Or equivalently
      \label{eq:xxie5}
      {L_2 \over c} > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)

@@

        (1 - \hat\phi) e^r + \hat\phi

@@

        p [(K + q)^{-\alpha} b + K^{-\alpha}] K^\alpha

@@

        \alpha \xi [(1 - \hat\phi) e^r + \hat\phi]^\xi

@@

    Meanwhile
      c &<& {p \over \xi}
        (1 - \hat\phi) e^r
        \alpha K^\alpha p \left[ (K + q)^{-\alpha - 1} b + K^{-\alpha - 1} \right]
        \xi (\alpha + 1) (1 - \hat\phi)^\xi e^{r \xi}
    Thus, to show inequality \eqref{eq:xxie5}, it is sufficient to show
      \label{eq:xxie6}
          b (K + q)^{-\alpha} + K^{-\alpha}
          (1 - \hat\phi) e^r + \hat\phi
        (\alpha + 1)
        \alpha^2
          b (K + q)^{-\alpha - 1} + K^{-\alpha - 1}
          (1 - \hat\phi) e^r
      } > {1 \over \alpha} \exp\left( {1 \over \alpha} \right)
    This simplies to
      \label{eq:xxie7}
      \hat\phi < {
        g - 1
        (1 - e^{-r}) g - 1
    where
    g = 
          (K + q)^{\alpha + 1} + b K^{\alpha + 1}
        \right] \alpha e^{1/\alpha}
        (\alpha + 1) K (K + q) 
          (K + q)^{\alpha} + b K^{\alpha}
    \right\}^{1/\xi}
    Assumption \eqref{eq:xxie1.1} ensures $g > e^r/(e^r - 1)$. Hence the
    right-hand-side of inequality \eqref{eq:xxie7} is larger than 1.
    So \eqref{eq:xxie7} is satisfied by all $\hat\phi$. Thus
    \eqref{eq:xxie6} holds and we conclude

  \item When $C < \delta v$ implies $X < q$ for some $q > 0$, we have
    \begin{eqnarray*}
      && \mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
      &=& \alpha p K^\alpha (1 + b) \int_{-\infty}^0
      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi}
      (K - x)^{-\alpha-1} dx \\
      &&
      + (1 - p) \beta K'^\beta \left[
        (1 + b) \int_0^q + \int_q^\infty
      \right]
      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi} (K' + x)^{-\beta-1} dx    
    \end{eqnarray*}
  \end{enumerate}
  So the same calculation as in the previous case carries through
  with $q = 0$. The same conclusion of
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  stands.
\end{proof}

@@

\bibliographystyle{unsrt}
\bibliography{my_citations}
\end{document}
